---
author: "Nicholas Spyrison"
date: 2019-01-24
linktitle: PDF Sense Manual tours
title: PDF Sense Manual tours
# categories: [ "Development", "hugo" ]
# tags: ["hugo", "theme", "html", "css"]
weight: 10
---
```{r, include=F}
library(spinifex)
# fig 7
load("../../data/jetsProj_sub.rda")
load("../../data/PDFSense_fig7_basis.rda")
# fig 8
load("../../data/grDIScenter.rda")
load("../../data/PDFSense_fig8l_basis.rda")
```

[PDS Sense paper - google scholar link](https://scholar.google.com/scholar?cluster=9110690189452049647&hl=en&as_sdt=0,5)



### Manual Tours from figure 7
```{r, warning=F}
# Fig 7, subset to ATLAS7old and ATLAS7new
view_basis(fig7_basis, labels = colnames(jetsProj_sub[, 1:4]))
play_manual_tour(data = jetsProj_sub[, 1:4], basis = fig7_basis, manip_var = 3,
                 cat_var = jetsProj_sub$exp)
play_manual_tour(data = jetsProj_sub[, 1:4], basis = fig7_basis, manip_var = 4,
                 cat_var = jetsProj_sub$exp)
```
Manipulating PC3 rotates variation of ATLAS7new out of the projection.

Manipulating PC4 rotates some variation of ATLAS7nold out of the projection.

### Manual Tours from figure 8
```{r, warning=F}
# Fig 8 left, _center data.
view_basis(fig8l_basis, labels = colnames(grDIScenter[, 1:6]))
play_manual_tour(data = grDIScenter[, 1:6], basis = fig8l_basis, manip_var = 6,
                 cat_var = grDIScenter$disID)
play_manual_tour(data = grDIScenter[, 1:6], basis = fig8l_basis, manip_var = 5,
                 cat_var = grDIScenter$disID)
```
manipulating PC6 rotates variation of 'dimuonSIDIS' experiements out of the projection.

manipulating PC5 rotates some of variation of 'DIS HERA1+2' experiements out of the projection.

#### Note that: 
- 4 plotly objects knits to 15MB .html file
- 2 plotly objects knits to 10MB .html file
- gganimate opens local .gif, should be able to make gif and include if needed
    - gganimate version current as of 24/01/2019